Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-04-06 ms.tex
conclusions 1

We have calculated steady-state models for the hot gaseous circumnuclear media of quiescent galaxies, under the assumption that gas is supplied exclusively by stellar wind mass loss and heated by shocked stellar winds, supernovae and black hole feedback.  We compute solutions for a range of different black hole masses, heating rates, and observationally-motivated stellar density profiles.  Our numerical results (Table \ref{table:models}) are used to verify and calibrate analytic relationships (Appendix \ref{app:rs}, \ref{app:be}).  The latter are in turn used to explore systematically how the SMBH accretion rate varies with black hole mass and the galaxy's star formation history.  Our results for $\dot{M}(\Mbh)$ are compared with observed trends of the nuclear X-ray luminosities of quiescent SMBHs and low luminosity AGN, and used to address other key questions, such as the thermal stability of their accretion flows.  Our conclusions may be summarized as follows.

  \begin{enumerate}
  %Younger population=>higher accretion rate; generally higher eta and
  %lower vw except for very young stellar populations.
  \item A stagnation radius, $\rs$, divides the nuclear gas between an accretion flow and an
    outward wind. In a steady state the gas inflow rate towards the black hole is proportional to the
    stellar mass enclosed inside of $\rs$.  In the limit of strong heating, the stagnation radius is close to the Bondi radius (eq.~[\ref{eq:rbondi}]).  
  \item The location of the stagnation radius, and thus the SMBH accretion rate, is sensitive to the effective heating rate of the gas.  Stellar wind heating, supernovae, and AGN feedback depend on the SMBH mass as well as the star
    formation history in the galactic nucleus.  A recent burst of star formation within the last $\sim 10$ Myr results in mass and energy injection both being dominated by fast stellar winds from young stars ($v_w \sim 1000$ km s$^{-1}$). If star formation is instead relegated to the distant past, mass input becomes dominated by old stars while heating is dominated by AGN feedback and supernovae.  For nuclei with ongoing, continuous star formation, energy input is usually dominated by the youngest stars, while mass input is dominated by the old stellar population (Appendix \ref{app:windheat}).

Quantitatively $\rs\propto v_w^{-2}$, which implies that $\Mdot\sim v_w^{-3.8 (-2.4)}$ for core (cusp) galaxies.   However, $\Mdot$ will increase even more rapidly with decreasing $v_w$ for two reasons: (1) when $v_w$ becomes comparable to the stellar velocity dispersion ($\zeta < \zeta_c$; eq.~$[\ref{eq:zetacrit}]$), gas remains bound to the stellar bulge and cannot flow out (Appendix \ref{app:be}).  (2) For sufficiently low $v_w$ the gas near the stagnation radius becomes thermally unstable, potentially resulting in cycles of alternating high and low accretion states ($\S\ref{sec:cycle}$; e.g.,\citealt{Ciotti+10}).

>>>>

We have calculated steady-state models for the hot gaseous circumnuclear media of quiescent galaxies, under the assumption that gas is supplied exclusively by stellar wind mass loss and heated by shocked stellar winds, supernovae and black hole feedback.  We compute solutions for a range of different black hole masses, heating rates, and observationally-motivated stellar density profiles.  Our numerical results (Table \ref{table:models}) are used to verify and calibrate analytic relationships (Appendix \ref{app:rs}, \ref{app:be}).  The latter are in turn used to explore systematically how the SMBH accretion rate varies with black hole mass and the galaxy's star formation history.  Our results for $\dot{M}(\Mbh)$ are compared with observed trends of the nuclear X-ray luminosities of quiescent SMBHs and low luminosity AGN, and used to address other key questions, such as the thermal stability of their accretion flows.  Our conclusions are summarized as follows.

  \begin{enumerate}
  \item A stagnation radius, $\rs$, divides the nuclear gas between an accretion flow and an
    outward wind. In steady-state the gas inflow rate towards the black hole is proportional to the
    stellar mass enclosed inside of $\rs$.  In the limit of strong heating, the stagnation radius resides close to the Bondi radius (eq.~[\ref{eq:rbondi}]).  
  \item The location of the stagnation radius, and hence the SMBH accretion rate, is a sensitive function of the gas heating rate $\propto v_w^{2}$.  Stellar wind heating, supernovae, and AGN feedback depend explicitely on the SMBH mass as well as the star formation history in the galactic nuclear region.  A recent star burst of age $\lesssim 10$ Myr produces mass and energy injection dominated by young stellar winds ($v_w \sim 1000$ km s$^{-1}$), while for an exclusively old stellar population, mass input is dominated by AGB winds and heating is dominated by AGN feedback and supernovae.  Ongoing, continuous star formation presents a hybrid situation: energy input is usually dominated by young stars, while mass input is dominated by the old stellar population (Appendix \ref{app:windheat}).

Quantitatively, $\rs\propto v_w^{-2}$, implying that $\Mdot\sim v_w^{-3.8 (-2.4)}$ for core (cusp) galaxies.  However, $\Mdot$ may increase even more rapidly as $v_w$ decreases for two reasons: (1) when $v_w$ is comparable or less than the stellar velocity dispersion ($\zeta < \zeta_c$; eq.~$[\ref{eq:zetacrit}]$), then gas remains bound to the stellar bulge and cannot produce a flow with a positive kinetic energy at infinity (Appendix \ref{app:be}); (2) For low $v_w$, gas near the stagnation radius becomes thermally unstable, which based on previous studies (e.g.,\citealt{Ciotti+10}) is found to result in time-dependent cycles of alternating high and low accretion states ($\S\ref{sec:cycle}$).
